### 2023-12-23 chapters/discussion.tex
Updates

\section{Representation learning with variational autoencoders}\label{sec_discussion:representation-learning-with-vaes}
%
In this thesis we studied two different approaches to speech representation learning: VAEs in \cref{chp:paper-hierarchical,chp:paper-modelagnostic,chp:paper-benchmarking} and self"=supervised methods in \cref{chp:paper-brief}. 
As we saw in \cref{chp:paper-hierarchical,chp:paper-modelagnostic}, the probabilistic formulation of VAEs provides benefits for their application to uncertainty quantification, although competitive methods based on self"=supervised foundation models have also been successful \parencite{xiao_we_2021,hendrycks_using_2019,bergman_classificationbased_2020}. 
Nonetheless, as discussed in \cref{chp:paper-brief}, self"=supervised methods are superior to VAEs when it comes to performance on most other downstream tasks, such as speech recognition and spoken language understanding. 
While this statement of course depends on the task and the amount and type of unlabeled and labeled data, self"=supervised methods for speech are generally evaluated on downstream tasks that are harder and more complex than those used to evaluate VAEs. This is evidenced by the leaderboards on benchmarks such as SUPERB, SLUE and ZeroSpeech \parencite{yang_superb_2021,shon_slue_2021,dunbar_zero_2021}.\footnote{One example of comparable evaluation is that between \cref{tab: phoneme recognition (PER),tab:unsupervisedASR} where self"=supervised methods can be seen to outperform VAEs for phoneme recognition while self"=supervised methods are also reported for word-level speech recognition.} 
In the following, we will discuss potential directions of future research and try to shed light on why representations from VAEs underperform on downstream tasks and how they might be improved.

@@

Including masked objectives into the principled probabilistic framework of VAEs might require rethinking the principles themselves, but it seems likely that relaxing the requirement of exact input reconstruction and enforcing the need for encoding (and decoding) contextual information could pave the way for future types of VAEs to learn representations that are competitive with self"=supervised methods.

\moretodo[inline]{Out-of-distribution detection on speech?}

>>>>

\section{Representation learning with variational autoencoders}\label{sec_discussion:representation-learning-with-vaes}
%
In this thesis we studied two different approaches to speech representation learning: VAEs in \cref{chp:paper-hierarchical,chp:paper-modelagnostic,chp:paper-benchmarking} and self"=supervised methods in \cref{chp:paper-brief}. 
As we saw in \cref{chp:paper-hierarchical,chp:paper-modelagnostic}, the probabilistic formulation of VAEs provides benefits for their application to uncertainty quantification, although competitive methods based on self"=supervised foundation models have also been successful \parencite{xiao_we_2021,hendrycks_using_2019,bergman_classificationbased_2020}. 
Nonetheless, as discussed in \cref{chp:paper-brief}, self"=supervised methods are superior to VAEs when it comes to performance on most other downstream tasks, such as speech recognition and spoken language understanding. 
While this statement of course depends on the task and the amount and type of unlabeled and labeled data, self"=supervised methods for speech are generally evaluated on downstream tasks that are harder and more complex than those used to evaluate VAEs. This is evidenced by the leaderboards on benchmarks such as SUPERB, SLUE and ZeroSpeech \parencite{yang_superb_2021,shon_slue_2021,dunbar_zero_2021}.\footnote{One example of comparable evaluation is that between \cref{tab: phoneme recognition (PER),tab:unsupervisedASR} where self"=supervised methods can be seen to outperform VAEs for phoneme recognition while self"=supervised methods are additionally reported for word-level speech recognition.} 
In the following, we will discuss potential directions of future research and try to shed light on why representations from VAEs underperform on downstream tasks and how they might be improved.

@@

Including masked objectives into the principled probabilistic framework of VAEs might require rethinking the principles themselves. Nonetheless, it seems likely that relaxing the requirement of exact input reconstruction and enforcing the need for encoding (and decoding) contextual information could pave the way for future types of VAEs to learn representations that are competitive with self"=supervised methods.

\section{Out-of-distribution detection for speech}
%
% \moretodo[inline]{Out-of-distribution detection on speech?
%
% Outline:
% Intro paragraph drawing from OOD detection on images.
% Differences and unique challenges. Generally a harder problem than OOD detection for images

The studies of \cref{part:unsupervised-uncertainty-estimation} dealt with OOD detection in the context of image classification, somewhat casually ignoring the fact that this thesis is mainly interested in audio and speech data. Image classification has been the de facto standard task for benchmarking OOD detection methods

For certain image distributions, it might be possible to deem an image OOD due to the color or intensity of a single pixel. 
However, for most complex distributions we are usually interested in a test relating to more semantic properties of the data - for instance the animal class, as discussed in the introduction (\cref{subsec:understanding-uncertainty}).

For speech, this property is similar; a single amplitude value is unlikely to provide much information about the distributional origin of a speech sample. Instead, whether a sample is OOD must be inferred from correlations and patterns found in amplitudes along the temporal dimension.

One study \parencite{bukhsh_outofdistribution_2023} ... (OOD detection on audio)

% - Temporal scale, can be OOD in local segments but not entire sequence. 
% - Harder to work with: Images provide immediate, human-interpretable feedback but audio requires playback which takes time.
%

### 2023-12-23 chapters/introduction.tex
Add some files
Fix decision-support -> decision support
Updates

\chapter[introduction]{Introduction}\label{chp:introduction}
%
% SIZE: ~9 pages
%
% OUTLINE: Broad introduction to the topic.
% - Motivate the reader.
% - Why should they care about this topic? 
% - What is the historical context of decision-support, what are the practical implications of it, and what happens when it fails.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{uber_nhsa_accident.png}
    \caption[A pedestrian was killed by an Uber self-driving car in Tempe, Arizona in 2018.]{ A pedestrian was killed by an Uber self-driving car in Tempe, Arizona in 2018. The car's sensors detected the pedestrian $5.6$ seconds before the crash, but the self-driving system failed to recognize its own uncertainty in multiple object misclassifications and so did not correctly predict her path or reduce the SUV's speed \parencite[photo credit]{nationaltransportationsafetyboardnhsa_collision_2019}.}
    % \caption{Aerial view of crash location showing path of pedestrian as she attempted to cross N. Mill Avenue and movement and speed of SUV at three points before impact. Pedestrian's path shows her position from initial detection ($5.6$ seconds before impact) until impact; SUV's position is shown at corresponding times beginning $4.2$ seconds before impact \parencite{nationaltransportationsafetyboardnhsa_collision_2019}.}
    \label{fig:uber_nhsa_accident}
\end{figure}

@@

One of the key advances needed to ensure that high-risk applications of machine learning comply with such requirements is accurate uncertainty estimation of their predictions. 
The multitude of rapid misclassifications made by the Uber self-driving car in Tempe is a likely indication that the model responsible for object-classification was overconfident in its predictions and unable to represent or communicate uncertainty to other systems in the car, or the safety driver. 
As a consequence, the model's many incoherent predictions were taken at face value and subsystems in the car acted on them without appropriate consideration to their reliability.
While any model will inevitably make wrong predictions, accurately estimating the associated uncertainty and communicating it to other systems or humans is necessary to reduce the risk that wrong predictions lead to catastrophic failures. 
Fittingly, a recent study by the European Parliamentary Research Services \parencite{europeanparliament_artificial_2022} concluded: 
%
\begin{quote}
    \centering\itshape
    Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions.
\end{quote}
% \vspace{0.5em}
\begin{center}
\noindent\rule{0.2\textwidth}{0.5pt}
\end{center}
\vspace{1em}
%
%
% OUTLINE: Primer overview of the thesis. What are the topics and what is the context of the research.
% \noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
\noindent The research presented in this thesis deals with the use of machine learning for medical interviews with a focus on uncertainty estimation and speech representation learning. 
The research project was partially funded by a grant from the national Danish Innovation Fund (grant no.\@ 0153-00167B) and defined and carried out in collaboration with the Danish health tech company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and to Corti and discuss the motivation behind the research project.

\section{Uncertainty by example: Corti use-cases}
%
% OUTLINE: Motivate the research project. 
% - What is the context of the research and why is it important? 
% - What are the current applications of machine learning in the domain (decision-support in healthcare)?
% - What are the practical implications of the research? 
% - What are the challenges and limitations of the current state-of-the-art?
%
Corti develops decision support software for healthcare professionals such as general practitioners, medical coders, and call-takers at emergency medical services. The system supports the professional similar to a copilot for medical interviews, for instance providing a graph-based protocol for triaging and helping to keep track of key information shared in the conversation.

>>>>

\chapter[introduction]{Introduction}\label{chp:introduction}
%
% SIZE: ~9 pages
%
% OUTLINE: Broad introduction to the topic.
% - Motivate the reader.
% - Why should they care about this topic? 
% - What is the historical context of decision support, what are the practical implications of it, and what happens when it fails.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=0.85\textwidth]{uber_nhsa_accident_cropped.png}
    \caption[A pedestrian was killed by an Uber self-driving car in Tempe, Arizona in 2018.]{ A pedestrian was killed by an Uber self-driving car in Tempe, Arizona in 2018. The car's sensors detected the pedestrian $5.6$ seconds before the crash, but the self-driving system failed to recognize its own uncertainty in multiple object misclassifications and so did not correctly predict her path or reduce the SUV's speed \parencite[photo credit]{nationaltransportationsafetyboardnhsa_collision_2019}.}
    % \caption{Aerial view of crash location showing path of pedestrian as she attempted to cross N. Mill Avenue and movement and speed of SUV at three points before impact. Pedestrian's path shows her position from initial detection ($5.6$ seconds before impact) until impact; SUV's position is shown at corresponding times beginning $4.2$ seconds before impact \parencite{nationaltransportationsafetyboardnhsa_collision_2019}.}
    \label{fig:uber_nhsa_accident}
\end{figure}

@@

One of the key advances needed to ensure that high-risk applications of machine learning comply with such requirements is accurate uncertainty estimation of their predictions. 
The multitude of rapid misclassifications made by the Uber self-driving car in Tempe is a likely indication that the model responsible for object-classification was overconfident in its predictions and unable to represent or communicate uncertainty to other systems in the car, or the safety driver. 
As a consequence, the model's many incoherent predictions were taken at face value and subsystems in the car acted on them without appropriate consideration to their reliability.
While any model will inevitably make wrong predictions, accurately estimating the associated uncertainty and communicating it to other systems or humans is necessary to reduce the risk that wrong predictions lead to catastrophic failures. 
Fittingly, a recent study by the European Parliamentary Research Services \parencite{europeanparliament_artificial_2022} concluded: 
%
\begin{quote}
    \centering\itshape
    Future AI solutions for healthcare should be implemented by integrating uncertainty estimation, a relatively new field of research that aims to provide clinicians with clinically useful indications on the degree of confidence in AI predictions.
\end{quote}
% \vspace{0.5em}
\begin{center}
\noindent\rule{0.2\textwidth}{0.5pt}
\end{center}
\vspace{1em}
%
%
% OUTLINE: Primer overview of the thesis. What are the topics and what is the context of the research.
% \noindent The research presented in this thesis will focus on the use of machine learning in relation to uncertainty estimation, speech processing, and medical conversations. 
\noindent The research presented in this thesis deals with the use of machine learning for medical interviews with a focus on uncertainty estimation and representation learning. 
The research project was partially funded by a grant from the national Danish Innovation Fund (grant no.\@ 0153-00167B) and defined and carried out in collaboration with the Danish health tech company Corti, which develops computer software for medical decision support. 
In the rest of this chapter, we will provide a high-level introduction to these topics and to Corti and discuss the motivation behind the research project.

\section{Uncertainty by example: Corti use-cases}
%
% OUTLINE: Motivate the research project. 
% - What is the context of the research and why is it important? 
% - What are the current applications of machine learning in the domain (decision support in healthcare)?
% - What are the practical implications of the research? 
% - What are the challenges and limitations of the current state-of-the-art?
%
Corti develops decision support software for healthcare professionals such as general practitioners, medical coders, and call-takers at emergency medical services. The system supports the professional similar to a copilot for medical interviews, for instance providing a graph-based protocol for triaging and helping to keep track of key information shared in the conversation.

### 2023-12-23 frontmatter/abstract-da.tex
Updates

Nylige fremskridt inden for naturlig sprogbehandling gør det muligt for maskinlæringsmodeller at deltage aktivt i medicinske interviews for at lette administrative byrder, forbedre dokumentation, og assistere i realtid.
Sundhedsplejen adskiller sig dog fra andre domæner på grund af den høje risiko forbundet med selv små fejl. Da ingen model er fejlfri, tilskynder dette til at associere modelprædiktioner med robuste usikkerhedsestimater, især i scenarier, der involverer ude-af-fordeling (OAF) data, såsom sjældne sygdomme.

Denne afhandling tager udgangspunkt i hovedhypotesen, at usuperviseret repræsentationslæring er nyttig til usikkerhedsestimering i medicinske opgaver.
Den giver følgende bidrag:
%
\begin{enumerate}[topsep=3pt, partopsep=0pt, itemsep=3pt, parsep=0pt, leftmargin=2em, label=(\alph*)] %, label=(\roman*)]
    \item En likelihood-ratio score til OAF-detektion med variationelle autoenkodere, der afhjælper den bevist negative effekt af lav-niveau features.
    \item En statistisk test til OAF-detektion, der kombinerer score- og typikalitetstests og kan bruges med likelihoods fra enhver differentiabel generativ model.
    \item Et benchmark for probabilistiske talerepræsentationsmodeller og en ny metode til at lære hierarkiske repræsentationer.
    \item En oversigt over usuperviseret repræsentationslæring til neural talebehandling og en tilsvarende modeltaksonomi.
    \item En fejlanalyse og revideret evaluering af state-of-the-art modeller til automatiseret medicinsk kodning på MIMIC-III og IV datasættene.
    \item Et retrospektivt studie af talebaseret genkendelse af stroke i præhospitale akuttelefonopkald der viser betydelig forbedring i forhold til opkaldstagere.
\end{enumerate}
%
Sammenfattende adresserer denne afhandling udfordringer indenfor usikkerhedsestimering og repræsentationslæring til tale og udforsker medicinske anvendelser af maskinlæring. 
Dens bidrag er afgørende i udviklingen af et operationelt beslutningsstøttesystem til medicinske interviews, der søger at øge kvaliteten af patientbehandlingen ved at understøtte effektiv, informeret beslutningstagen.

>>>>

Nylige fremskridt inden for naturlig sprogbehandling gør det muligt for maskinlæringsmodeller at deltage aktivt i medicinske interviews for at lette administrative byrder, forbedre dokumentation, og assistere i realtid.
Sundhedsplejen adskiller sig dog fra andre domæner på grund af den høje risiko forbundet med selv små fejl. Da ingen model er fejlfri, tilskynder dette til at associere modelprædiktioner med robuste usikkerhedsestimater, især i scenarier, der involverer ude-af-fordeling (UAF) data, såsom sjældne sygdomme.

Denne afhandling tager udgangspunkt i hovedhypotesen, at usuperviseret repræsentationslæring er nyttig til usikkerhedsestimering i medicinske opgaver.
Den giver følgende bidrag:
%
\begin{enumerate}[topsep=3pt, partopsep=0pt, itemsep=3pt, parsep=0pt, leftmargin=2em, label=(\alph*)] %, label=(\roman*)]
    \item En likelihood-ratio score til UAF-detektion med variationelle autoenkodere, der afhjælper den bevist negative effekt af lav-niveau features.
    \item En statistisk test til UAF-detektion, der kombinerer score- og typikalitetstests og kan bruges med likelihoods fra enhver differentiabel generativ model.
    \item Et benchmark for probabilistiske talerepræsentationsmodeller og en ny metode til at lære hierarkiske repræsentationer.
    \item En oversigt over usuperviseret repræsentationslæring til neural talebehandling og en tilsvarende modeltaksonomi.
    \item En fejlanalyse og revideret evaluering af state-of-the-art modeller til automatiseret medicinsk kodning på MIMIC-III og IV datasættene.
    \item Et retrospektivt studie af talebaseret genkendelse af stroke i præhospitale akuttelefonopkald der viser betydelig forbedring i forhold til opkaldstagere.
\end{enumerate}
%

Sammenfattende adresserer denne afhandling udfordringer indenfor usikkerhedsestimering og repræsentationslæring til tale og udforsker medicinske anvendelser af maskinlæring. 
Dens bidrag er afgørende i udviklingen af et operationelt beslutningsstøttesystem til medicinske interviews, der søger at øge kvaliteten af patientbehandlingen ved at understøtte effektiv, informeret beslutningstagen.
